---
title: "Functionazling Rasters From NEON"
author: "Jeff Atkins"
date: "June 21, 2016"
output: html_document
---


Start by putting up objectives/tasks that students will be working though:

1. Import a raster — A lidar canopy height model (lidar/Teak_lidarCHM.tif)
1. For the CHM, set values == 0 to NA (not trees)
1. classify the raster according to some distribution – low medium and tall trees. This could be done using a histogram potentially or we could just decide that <2m is generally grasses / understory, <6m small trees,and the rest are tall trees. A function could import the desired thresholds. Visualize histogram/density and plot vertical cutoff lines.

1. PLOT - layer the classified raster on top of the hillshade, add a legend for each “class” - legends are super tricky to simplifying this process with a function would be good.  see: http://neon-workwithdata.github.io/neon-data-institute-2016/R/classify-by-threshold-R/  for my take on forcing a legend outside of the plot area using par settings. You may have other better forms of magic to make this work well. :)

1. Export the plot figure to a pdf – publishable
1. Export the classified raster as a geotiff with NaFlagg = -9999 to an outputs folder.

```{r set-library }
library(raster)
library(rgdal)

```

## Step One: Import Raster

```{r import-raster}
chm <- raster("../NEONdata/D17-California/TEAK/2013/lidar/Teak_lidarCHM.tif")
```

## Step Two:  Set values == 0 to NA (not trees)  

```{r set-values-NA}
#Sets those 0s to NA
chm[ chm <= 0 ] <- NA

#check this raster out, does it look like trees?
plot(chm,
     main="Canopy Height \n LowerTeakettle, California")
```

## Step Three:  Raster Classification By Height

```{r raster-height-classify}
#lets do a histogram and get some distribution
hist(chm)

#Want to focus on tree size class
height.class.m <- c(0, 15, 1,
                         15, 30, 2,
                         30, 75, 3)

reclass.m <- matrix(height.class.m,
                    ncol = 3,
                    byrow = TRUE)


chm.class <- reclassify(chm, reclass.m)

#here is a draft plot by class
plot(chm.class,
     main = "Teakettle Canopy Height By Class")
```

## Step Four:  I Am Legend
## STEP Five:  Export to .pdf!

```{r plot-with-legend}

# plot outside of the plot region

# make room for a legend
par(xpd = FALSE, mar=c(5.1, 4.1, 4.1, 4.5))

# plot
pdf("../outputs/chm.class.pdf")
p.chm <- plot(chm.class,
     col=c("grey","blue","green"), # hard code colors, small trees (1)=white,
		 #medium (2) = blue,  large trees (3) = green
     main="Tree Canopy Height \nLower Teakettle",
     legend=F)
# allow legend to plot outside of bounds
par(xpd=TRUE)

# create the legend
legend((par()$usr[2] + 20), 4103300,  # set x,y legend location
       legend = c("Small", "Medium", "Large"),  # make sure the order matches the colors, next
       fill = c("grey", "blue", "green"),
       bty="n") # turn off border
dev.off()
```

## Step Six:  Export the raster

```{r export-raster}
writeRaster(chm.class,
            filename="../outputs/TEAK/Teak_nsAspect.tif",
            format="GTiff",
            options="COMPRESS=LZW",
            overwrite = TRUE,
            NAflag = -9999)


```