---
title: "Functionazling Rasters From NEON"
author: "Jeff Atkins"
date: "June 21, 2016"
output: html_document
---


Start by putting up objectives/tasks that students will be working though:

1. Import a raster — A lidar canopy height model (lidar/Teak_lidarCHM.tif)
1. For the CHM, set values == 0 to NA (not trees)
1. classify the raster according to some distribution – low medium and tall trees. This could be done using a histogram potentially or we could just decide that <2m is generally grasses / understory, <6m small trees,and the rest are tall trees. A function could import the desired thresholds. Visualize histogram/density and plot vertical cutoff lines.

1. PLOT - layer the classified raster on top of the hillshade, add a legend for each “class” - legends are super tricky to simplifying this process with a function would be good.  see: http://neon-workwithdata.github.io/neon-data-institute-2016/R/classify-by-threshold-R/  for my take on forcing a legend outside of the plot area using par settings. You may have other better forms of magic to make this work well. :)

1. Export the plot figure to a pdf – publishable
1. Export the classified raster as a geotiff with NaFlagg = -9999 to an outputs folder.

```{r set-library }
library(raster)
library(rgdal)
library(magrittr)

source("scripts/functions.R")
```

## Step One: Import Raster

```{r import-raster}

#### PARAMETERS
chm_file <-("../NEONdata/D17-California/TEAK/2013/lidar/Teak_lidarCHM.tif")

# Threshold breaks for canopy height in (m)
chosen.breaks <- c(6, 30, 50, 100)



chm <- raster(chm_file)

#quick sanity check
plot(chm,
     main="Canopy Height \n LowerTeakettle, California")
```

## Step Two:  Set values == 0 to NA (not trees)  

```{r set-values-NA}
#Sets those 0s to NA
chm[ chm <= 0 ] <- NA



density_plot_from_chm(chm, "Teakettle!", chosen.breaks)
hist_plot_from_chm(chm, "Teakettle Histogram", chosen.breaks)




make_pdf(density_plot_from_chm(chm, 
                               title = "Canopy heights for NEON TeaKettle",
          chosen.breaks),
         filename = "teak_chm_density.pdf", width = 8, height = 6)   

make_pdf(hist_plot_from_chm(chm, 
                               title = "Canopy heights for NEON TeaKettle",
                               chosen.breaks),
         filename = "teak_chm_hist.pdf", width = 8, height = 6)
         
reclassify.mat <- create_height_class_matrix(chosen.breaks)
reclassified.chm <- reclassify(chm, reclassify.mat)

#plotting reclassified chm
make_pdf(plot_reclassified_raster(reclassified.chm,
                                  "Teakettle", 
                                  chosen.breaks),
         filename = "teakettle_reclassified.pdf")
```

```{r export-raster}
# writeRaster(chm.class,
#             filename="outputs/Teak_nsAspect.tif",
#             format="GTiff",
#             options="COMPRESS=LZW",
#             overwrite = TRUE,
#             NAflag = -9999)
# 
# 
# ####This creates a great log session text file at the end!!!!!!
# sink(paste0(format(Sys.time(), "%Y-%m-%d_%H%M%S"),
# "_sessionInfo.txt"))     #zero after paste has no spaces
# sessionInfo()
# sink()
```